Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-18 p2/ordering.tex
Style tweaks
Improvements
Further minor improvements

As we recognize that multiple orders may be possible,
let us describe our ordering algorithm by
defining a subset of the permutations, $\ordersetsign$.
As we will see, $\ordersetsign$ contains all possible
sequences of commands that do not break all filesystems.
In \cref{simple_reorder_equiv}, we offer a formal proof of this.
The facts that all valid reorderings of simple sequences are equivalent,
and all equivalent sequences are reorderings, are important 
properties of simple sequences.

We define $\ordersetsign$ with the help of a partial order.
\begin{mydef}[$\orderrel$]\mlabel{def_orderrel}
The binary relation $\orderrel$ holds for some command pairs
on directly related nodes in the following cases:
\begin{itemize}
\item $\caaa{x}{y}{\parentf{n}} \orderrel \caaa{z}{w}{n}$ if both commands are construction commands, and
\item $\caaa{x}{y}{n} \orderrel \caaa{z}{w}{\parentf{n}}$ if both commands are destruction commands.
\end{itemize}
\end{mydef}
This relation is clearly irreflexive and anti-symmetric,
and as a command cannot be both a construction and a destruction command and
the relation follows the tree of nodes, it is acyclic,
and thus defines a partial order.

\begin{mydef}[$\ordersetsign$]\mlabel{def_orderset}
For a simple sequence of commands $\wrksx{S}$,
$\orderset{S}$ is the set of permutations satisfying the partial order defined by $\orderrel$.
% For a simple set or sequence of commands $S$,
% we use $\orderset{S}$ to denote the set of 
% permutations of the commands in $S$
% for which the following holds:
% for any subsequence $\cxynv\cc\czwmv$,
% \begin{itemize}
% \item if $n=\parentf{m}$, then $y=\vald$,
% \item if $\parentf{n}=m$, then $y=\empt$.
% \end{itemize}
\end{mydef}

% As usual, by subsequence we mean a subset of the commands in $S$ in the same order as they are found in $S$.
The simplicity of the definitions mean that sorting the commands
and enumerating
all sequences in $\ordersetsign$ can be implemented by a straightforward
algorithm.

The principal ideas behind the ordering are 
\cref{ax_separate_commute}, from which we know that commands on incomparable
nodes can be applied in any order, so we only need to focus on the order of commands
with comparable nodes;
\cref{connected_changes}, from which we know that if a simple sequence contains commands on two comparable
nodes, then it contains commands on all nodes in between, so it is enough to specify
the order of commands on directly related nodes;
and finally, \cref{simple_distant_pairs}, from which we know
that commands on parent--child pairs must be construction or destruction pairs.
% Investigating these pairs we see that they are identified by the output value of the first command ($y$),
% which in turn determines whether the command on the parent or the child must precede the other.

@@

@@

If $\cxynv$ is a replacement command, then we know $[x]=[y]=[\valfx]$, and
$[\FS(n)]=[\valfx]$. If $\FS(n)=y$, then instead of this filesystem,
consider $\fsreplacement{\FS}{n}{\valf}$ where $\valf$ is any file value
other than $y$.
(We assume that there are more than one file values; if this is not so,
$\cffa{n}$ becomes an assertion command.)
From here the proof concludes in the same way as above.
\end{proof}

>>>>

As multiple orders may be possible,
we describe our ordering algorithm by
defining a subset of the permutations of the commands, $\ordersetsign$.
\Cref{simple_reorder_equiv} proves that $\ordersetsign$ indeed contains all possible
sequences of commands that do not break all filesystems.
The facts that all valid reorderings of simple sequences are equivalent,
and all equivalent sequences are reorderings follow from this \namecref{simple_reorder_equiv}, and are important 
properties of simple sequences.

We define $\ordersetsign$ with the help of a partial order.
\begin{mydef}[$\orderrel$]\mlabel{def_orderrel}
The binary relation $\orderrel$ holds for some command pairs
on directly related nodes in the following cases:
\begin{itemize}
\item $\caaa{x}{y}{\parentf{n}} \orderrel \caaa{z}{w}{n}$ if both commands are construction commands, and
\item $\caaa{x}{y}{n} \orderrel \caaa{z}{w}{\parentf{n}}$ if both commands are destruction commands. \qedhere
\end{itemize}
\end{mydef}
This relation is clearly irreflexive and anti-symmetric,
and as a command cannot be both a construction and a destruction command and
the relation follows the tree of nodes, it is acyclic.
Therefore its transitive closure, also denoted by $\orderrel$, is a partial order.

\begin{mydef}[$\ordersetsign$]\mlabel{def_orderset}
For a simple sequence of commands $\wrksx{S}$,
$\orderset{S}$ is the set of permutations of the commands in $S$ respecting the partial order defined by $\orderrel$,
that is,
for any $T\in\orderset{S}$ and any $\alpha,\beta\in T$, if $\alpha\orderrel\beta$ then $\alpha$
precedes $\beta$ in $T$.
% For a simple set or sequence of commands $S$,
% we use $\orderset{S}$ to denote the set of 
% permutations of the commands in $S$
% for which the following holds:
% for any subsequence $\cxynv\cc\czwmv$,
% \begin{itemize}
% \item if $n=\parentf{m}$, then $y=\vald$,
% \item if $\parentf{n}=m$, then $y=\empt$.
% \end{itemize}
\end{mydef}

% As usual, by subsequence we mean a subset of the commands in $S$ in the same order as they are found in $S$.
Returning a valid $T\in\orderset{S}$ order for a set of commands $S$
can be implemented using any well-known algorithm for topological sorting,
where the directed acyclic graph that is the input of the algorithm is defined by $\orderrel$.

The principal ideas behind the definition of $\orderrel$ are 
\cref{ax_separate_commute}, from which we know that commands on incomparable
nodes can be applied in any order, so we only need to focus on the order of commands
with comparable nodes;
\cref{connected_changes}, from which we know that if a simple sequence contains commands on two comparable
nodes, then it contains commands on all nodes in between, so it is enough to specify
the order of commands on directly related nodes;
and finally, \cref{simple_distant_pairs}, from which we know
that commands on parent--child pairs must be construction or destruction pairs.
% Investigating these pairs we see that they are identified by the output value of the first command ($y$),
% which in turn determines whether the command on the parent or the child must precede the other.

@@

This follows directly from \cref{def_orderset}.

@@

If $\cxynv$ is a replacement command, then we know $[x]=[y]=[\valfx]$, and
$[\FS(n)]=[\valfx]$. If $\FS(n)=y$, then instead of this filesystem,
consider $\fsreplacement{\FS}{n}{\valf}$ where $\valf$ is any file value
other than $y$.
From here the proof concludes in the same way as above.
\end{proof}

### 2018-07-18 p2/upd-def.tex
Further minor improvements

In a command-based reconciliation solution we assume that we have two sequences of commands
$A$ and $B$ that have modified a single filesystem $\FS$ 
yielding two different replicas $\FS_1$ and $\FS_2$ which we
need to reconcile. While it is conceivable that the sequences are based on a record of
all operations that modified the two filesystems, in most filesystem implementations
such records are not implemented, and therefore we must construct suitable sequences
by comparing $\FS_1$ (and $\FS_2$) to their common ancestor, $\FS$. 
This is called \emph{update detection.}

The set of commands $U$ necessary to transform $\FS$ into $\FS_1$ can be collected
by inspecting the (necessarily finite) union of non-empty nodes in the two filesystems.
If at node $n$, $\FS(n)=x$ and $\FS_1(n)=y\neq x$, then we add the command $\cxyaa{n}$ to $U$.
We can always do so as there is a command available for all combinations of input and output types and values.
We see that any suitable $U$ necessarily contains commands on all nodes at which the values have changed.

To be able to reason about sets and sequences of commands, let us introduce the following properties:

\begin{mydef}[Minimal and simple command sets and command sequences]
A sequence or set of commands is \emph{minimal} if it contains at most one command on each node.
It is \emph{simple} if it is minimal and it does not contain assertion commands.
\end{mydef}

>>>>

In a command-based reconciliation solution we assume that we have two sequences of commands
$A$ and $B$ that have modified a single filesystem $\FS$ 
yielding two different replicas $\FS_1$ and $\FS_2$ which we
need to reconcile. While it is conceivable that the sequences are based on a record of
all operations that modified the two filesystems, in most filesystem implementations
such records do not exist, and therefore we must construct suitable sequences
by comparing $\FS_1$ (and $\FS_2$) to their common ancestor, $\FS$. 
This is called \emph{update detection.}

The set of commands $U$ necessary to transform $\FS$ into $\FS_1$ can be collected
by inspecting the (finite) union of non-empty nodes in the two filesystems.
If at node $n$, $\FS(n)=x$ and $\FS_1(n)=y\neq x$, then we add the command $\cxyaa{n}$ to $U$.
We can always do so as there is a command available for all combinations of input and output types and values.
We see that any suitable $U$ necessarily contains commands on all nodes at which the values have changed.

% To be able to reason about sets and sequences of commands, let us introduce the following properties:

\begin{mydef}[Minimal and simple command sets and sequences]\mlabel{def_min_simp}
A sequence or set of commands is \emph{minimal} if it contains at most one command on each node.
It is \emph{simple} if it is minimal and it does not contain assertion commands.
\end{mydef}
